% DMIR x microNAS — working draft
% Target: IEEE Signal Processing Letters (IEEEtran journal mode).
% Rules for this file:
%   * every number must come from logs/experiments.jsonl or a cited source;
%   * placeholders are \todo{...} — none may survive to submission;
%   * check prose against paper/STYLE.md before committing.
\documentclass[journal]{IEEEtran}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}

\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

\begin{document}

\title{Hardware-Constrained Architecture Search for Lane-Change Intention
Prediction from Vehicle Time Series}
% Alternative titles, pick later:
% - Microcontroller-Ready Lane-Change Intention Prediction via Constrained NAS
% - Searching Small: NAS under MCU Budgets for Driving Intention Time Series

\author{Sepehr~Mohammady,
        \todo{co-authors, order},
        \todo{supervisor}%
\thanks{The authors are with the ELIOS Lab, Department of Electrical,
Electronic and Telecommunications Engineering and Naval Architecture (DITEN),
University of Genoa, Genoa, Italy.}%
\thanks{This work was supported by the SYNERGIES project
\todo{grant number and official acknowledgment text}.}}

\maketitle

\begin{abstract}
Predicting a driver's lane-change intention early enough to act on it is a
time-series classification problem with hard deployment limits: the model has
to run on automotive-grade microcontrollers within tight memory and latency
budgets. We apply constrained neural architecture search to the Lane Change
Intention Recognition driving dataset \cite{lcdataset} (windows of 50 samples
over 31 signals) and obtain compact one-dimen-
sional convolutional networks for three tasks: three-class intention
classification and time-to-lane-change regression for left and right
maneuvers. \todo{results: accuracy, RMSE, model size, latency on STM32}
The searched models improve on the published baseline (Transformer,
0.51~s time-to-lane-change RMSE, deployed in FP32 \cite{lcintention2026})
while using \todo{X}$\times$ less memory, and we report measured, not
estimated, footprint and latency on the same STM32 targets after int8
quantization.
\end{abstract}

\begin{IEEEkeywords}
Neural architecture search, lane-change intention, time-series classification,
TinyML, embedded inference, STM32.
\end{IEEEkeywords}

\section{Introduction}
\todo{Motivation: intention prediction as advance warning for ADAS; why
on-device inference (latency, privacy, cost, no connectivity assumptions);
the gap: published DMIR models target accuracy, not deployment budgets.}

Contributions:
\begin{enumerate}
  \item a constrained search space and objective tuned to one-dimensional
        driving signals under microcontroller budgets;
  \item models that beat the published DMIR baseline on all three tasks at a
        fraction of its footprint \todo{verify once results exist};
  \item measured deployment numbers (flash, RAM, latency) on
        \todo{STM32 part}, with the full pipeline released as open source.
\end{enumerate}

\section{Related Work}
\subsection{Lane-change intention prediction}
\todo{The LC dataset and framework baseline \cite{lcintention2026} (TTLC
regression, Transformer RMSE 0.51~s, FP32 deployment on STM32H7B3/F401);
earlier HighD-based TTLC work (RMSE 0.63--0.7); MicroNAS for time series and
TinyTNAS as closest NAS relatives.}

\subsection{Constrained architecture search for microcontrollers}
\todo{muNAS \cite{liberis2021munas}, MCUNet/TinyNAS, MicroNets; what transfers
to 1D time series and what does not.}

\section{Data}
We use the Lane Change Intention Recognition dataset \cite{lcdataset}
(50 human drivers in the CARLA simulator, 10~Hz, over 3{,}400 annotated lane
changes; MIT license), prepared as fixed windows of $T{=}50$ samples
(5~s) over $F{=}31$ channels \todo{confirm channel order and whether channel
31 is fileTime — drop it if so; official feature count is 30}. Classification uses balanced splits of 94{,}620 / 16{,}332 /
19{,}722 windows (train/val/test) over three classes: no intention, right
lane change, left lane change. The regression tasks map a window to the time
until the lane change, $t \in [0, 4]$~s discretized at $0.1$~s, with
39{,}313 / 7{,}742 / 8{,}097 (LCL) and 33{,}201 / 5{,}828 / 7{,}073 (LCR)
windows.

The splits are user-independent: we verified by matching raw per-driver
session logs against the prepared windows (correlation over 50-sample
windows) that the validation and test drivers follow the protocol of
\cite{lcintention2026} with no window leakage across splits.
The test splits contain isolated spikes (up to $5\times10^{6}$) on the
right/left \texttt{curvatureDx} channel pair — a numerical-derivative
artefact absent from training; we clip validation and test features to the
per-channel training range and leave training data untouched.

\section{Method}
\subsection{Search space}
The search space is built from one-dimensional convolutional blocks: standard,
depthwise and $1\times1$ convolutions with kernel sizes in $\{3,5,7\}$ and
channel counts up to 128, optional stride-2 and pooling, blocks connected
serially or with residual branches, and a head of one to three
fully-connected layers. The operator set is restricted to layers the ST Edge AI
compiler maps efficiently. \todo{state the exact ranges once frozen.}

\subsection{Constraints and search algorithm}
We use aging evolution: a population of architectures is sampled, the fittest of
a random subset is mutated by a network morphism, its child is trained, and the
oldest member is retired. Fitness combines four objectives---validation error,
peak RAM, model size and multiply--accumulate count---by the random
scalarisation of \cite{liberis2021munas}: each is normalised by a user budget
and divided by a random weight resampled every round, so a candidate exceeding
any budget is penalised first. Budgets were set from the STM32H7B3I-DK.
\todo{state population/sample sizes, rounds, and search cost.}

\subsection{Discovered architecture}
Across tasks the search independently arrives at the same microcontroller-friendly
recipe: $1\times1$ convolutions to re-shape the channel dimension, depthwise
convolutions for temporal structure, aggressive early downsampling, and a small
flattened head (98--308 features against the reference's 6400). Topology,
however, follows the task: the classifiers are essentially sequential, whereas
the regression model uses two parallel branches merged by an addition---the
search space permits branches everywhere, but only regression exploited them.
The classification model reduces the sequence $50 \rightarrow 25
\rightarrow 13 \rightarrow 7 \rightarrow 4$ before its head:

\begin{center}\small
\texttt{50$\times$31} $\rightarrow$ MaxPool $\rightarrow$ DWConv(s2)
$\rightarrow$ Conv$_{1\times1}$(77) $\rightarrow$ DWConv $\rightarrow$
DWConv(s2) $\rightarrow$ Conv$_{1\times1}$ $\rightarrow$ Add $\rightarrow$
AvgPool $\rightarrow$ FC(223) $\rightarrow$ FC(23) $\rightarrow$ FC(3)
\end{center}

This explains the footprint gap quantitatively. Both our model and the reference
CNN spend most of their flash in the fully-connected head, so the gain does not
come from avoiding a large head; it comes from what enters it. The reference
holds full temporal resolution and flattens a $50\times128$ map into 6400
features, giving a $64\times6400$ layer of 1.64~MB ($\sim$95\% of the model).
Ours pools first and flattens only $4\times77 = 308$ features, so it affords a
\emph{wider} head (223 units) in a $\sim$280~KB layer---a $6\times$ reduction.
The same downsampling cuts the convolutional work, yielding the $12.4\times$
fewer MACs and the $9.2\times$ measured speed-up of Table~\ref{tab:deploy}.

\section{Experiments}
\subsection{Setup}
\todo{training schedule, hardware, seeds/repeats, hyperparameters; all runs
logged to the released experiments.jsonl.}

\subsection{Results}
Table~\ref{tab:results} reports test-set metrics. All searched-model numbers
are our own test-set evaluation of the saved models, from searches run to their
full 150-round budget. \todo{add peak-RAM, MACs, and measured on-device latency
columns from ST Edge AI; consider re-running with an all-models save policy for
a guaranteed-optimal front (see notes).}

\begin{table}[t]
\caption{Test-set comparison. The internal reference models (unpublished) were
evaluated on the same test set; classification accuracy is our own evaluation,
the Transformer RMSE values are as reported. Size is the parameter count.}
\label{tab:results}
\centering
\begin{tabular}{lcccc}
\toprule
Model & Params & Acc.\ (\%) & LCR & LCL \\
\midrule
Reference CNN (cls) & 441\,k & 91.5 & --- & --- \\
Reference Transf.\ (reg, RMSE) & 333/49\,k & --- & 0.42 & 0.44 \\
Our DSCNN baseline & 10\,k & 91.5 & 0.32$^\dagger$ & 0.33$^\dagger$ \\
\midrule
Ours searched (cls / reg MAE) & 84\,k & \textbf{92.1} & 0.29$^\dagger$ & 0.33$^\dagger$ \\
Ours compact & 8\,k & 91.3 & --- & --- \\
Ours, no turn-signal & 21\,k & 91.1 & --- & --- \\
\bottomrule
\end{tabular}
\\[2pt]
\footnotesize $^\dagger$MAE (our search optimizes MAE); reference regression is
reported as RMSE. Our LCR RMSE is 0.45 and LCL 0.50.
\end{table}

For classification the searched model is both smaller and more accurate than the
441\,k reference CNN (92.1\% at 84\,k parameters; a 8\,k model still reaches
91.3\%). Regression beats the published single-TTLC SOTA (RMSE 0.51) at 2--3$\times$
fewer parameters (our RMSE 0.45 LCR, 0.47 LCL). We do not claim to beat the
internal-reference Transformers on regression RMSE (0.42/0.44), which remain
competitive small models; an RMSE-aware search narrowed but did not close that
gap. Removing the turn-signal channels drops classification
accuracy by about one point (92.1\% to 91.1\%)---evidence that the model
anticipates from vehicle dynamics and surrounding traffic rather than reading a
declared signal.
\todo{RMSE columns and the internal-reference RMSE (0.42/0.44) are not yet
matched because the search optimizes MAE; decide RMSE-aware objective vs
MAE-primary framing.}

\subsection{Deployment on STM32}
Target platform: STM32H7B3I-DK discovery kit (Cortex-M7 at 280~MHz, 1.4~MB
SRAM, 2~MB flash) — the same MCU family used by the baseline framework
\cite{lcintention2026}, which deployed FP32 models only. We therefore compare in
FP32, like for like, benchmarking both models on the same board through the ST
Edge AI Developer Cloud (Core 4.0.1, \texttt{balanced} optimization).
Table~\ref{tab:deploy} reports measured, not estimated, figures.

\begin{table}[t]
\caption{Measured on-device comparison, STM32H7B3I-DK (Cortex-M7, 280~MHz),
ST Edge AI Core 4.0.1, FP32.}
\label{tab:deploy}
\centering
\begin{tabular}{lccc}
\toprule
 & Reference CNN & Ours & Ours (tiny) \\
\midrule
Parameters      & 441\,k & 84\,k & 8\,k \\
Accuracy (\%)   & 91.7  & \textbf{92.1} & 91.3 \\
Latency (ms)    & 33.52 & 3.63 & \textbf{0.79} \\
Flash (KB)      & 1729  & 335  & \textbf{37} \\
RAM (KB)        & 38.3  & 9.2  & 9.2 \\
MACC (M)        & 1.97  & 0.16 & \textbf{0.03} \\
\bottomrule
\end{tabular}
\end{table}

The searched model is $9.2\times$ faster and $5.2\times$ smaller than the
reference CNN while being more accurate; a second operating point on the same
front gives up 0.4 points of accuracy for a $42\times$ speed-up and a
$47\times$ smaller model that infers in under a millisecond.

The per-layer breakdown explains why. Both our classifier and the reference spend
most of their flash in the fully-connected head, so the gain is not from
avoiding a large head but from what enters it: the reference flattens a
$50\times128$ map into 6400 features, giving a $64\times6400$ layer of 1.64~MB
(93\% of its weights), whereas the searched models pool first and flatten only
308 (or 98) features, cutting that layer $5.9\times$.

Across the four searched models this reliance falls monotonically with how much
of the search space the task exploits. The largest fully-connected layer holds
93\% of weights in the reference and 90.9\% in our tiny classifier, but 39.6\% in
the LCR model and only 11.4\% in the LCL model---whose largest layer is a
convolution. The same ordering appears in topology: both classifiers are
essentially sequential, whereas the LCR and LCL models fan out into two and three
parallel branches respectively. The search space permits branches for every task;
only regression used them. Topology follows the task, discovered rather than
imposed.

RAM behaves differently and is worth stating plainly, because three distinct
regimes appear across our models. Both searched classifiers measure
$\sim$9.2~KB regardless of a $9\times$ parameter gap: the $50\times31$ float32
input tensor alone occupies 6.2~KB and no architecture can go below it, so they
are \emph{input-bound}. The LCR model measures 20.8~KB and is \emph{width-bound}:
its widest branch emits a $25\times116$ tensor of 11.6~KB, and the peak working
set is that plus the input. The LCL model measures 26.5~KB and is neither---it is
\emph{liveness-bound}. Its widest activation (13.0~KB) and the next-widest
(12.0~KB) have disjoint live ranges and are never co-resident, so no
width-based rule reaches the measurement; instead its three-way merge pins the
6.2~KB input and two 3.3~KB branch outputs across many operators, giving a
24.8~KB peak, with the remainder attributable to allocator fragmentation. The
general statement is therefore that peak RAM is the maximum over the schedule of
all simultaneously live tensors, which collapses to
$\max(\text{input},\ \text{widest in\,+\,out})$ only for a chain. This is why the
RAM advantage ($\sim$4$\times$) is far smaller than the flash advantage.

On quantization, the ST toolchain accepts only float32/int8/uint8 tensors. Full
int8 shrinks the classifier to 104~KB and 1.885~ms but costs 5.2 accuracy points
on these wide-dynamic-range inputs ($92.1\% \rightarrow 86.9\%$; LCR MAE
$0.287 \rightarrow 0.449$), so we report float32 as the headline
configuration---which also matches the baseline's own FP32 deployment. An
int16$\times$8 scheme (int8 weights, int16 activations) preserves accuracy almost
exactly offline ($92.1\% \rightarrow 92.2\%$; LCR MAE
$0.287 \rightarrow 0.286$), but the toolchain does not deploy it. We verified
this on hardware rather than trusting the documentation: a 117.9~KB
int16$\times$8 file returns 326.15~KB of weights, byte-identical to the float32
build, while a genuinely int8 control compresses to 83.28~KB. The scheme is
silently dequantized to float32, so we report it as an offline bound only. The
episode is worth one sentence of method: a toolchain that accepts an unsupported
scheme and quietly widens it will produce a benchmark that looks like a result.
\todo{cite ST Edge AI Core 4.0 documentation (quantization / supported ops,
retrieved 2026-07-14).}
\todo{cls\_tiny on NUCLEO-F401RE still pending; see docs/research/deployment.md.}

\section{Conclusion}
\todo{two or three sentences; no new claims.}

\bibliographystyle{IEEEtran}
\begin{thebibliography}{9}
\bibitem{lcintention2026} L.~Forneris \emph{et al.}, ``A deployment-oriented
simulation framework for deep learning-based lane change prediction,''
\emph{IEEE Signal Process.\ Lett.}, vol.~33, pp.~136--140, 2026,
doi: 10.1109/LSP.2025.3638676.
\bibitem{liberis2021munas} E.~Liberis, {\L}.~Dudziak, and N.~D.~Lane,
``$\mu$NAS: Constrained neural architecture search for microcontrollers,'' in
\emph{Proc.\ 1st Workshop Mach.\ Learn.\ Syst.\ (EuroMLSys)}, 2021, pp. 70--79.
\bibitem{lcdataset} L.~Forneris \emph{et al.}, ``Lane change intention
recognition dataset,'' Zenodo, 2025, doi: 10.5281/zenodo.16686054.
\bibitem{forneris2024dmir} L.~Forneris \emph{et al.}, ``Setting up a
lightweight simulation environment for automated driving dataset
collection,'' in \emph{Applications in Electronics Pervading Industry,
Environment and Society (ApplePies)}, Springer LNEE, 2024, pp. 156--163,
doi: 10.1007/978-3-031-84100-2\_19.
\end{thebibliography}

\end{document}
